---
title: "Overview of the tidySE package"
author: "Stefano Mangiola"
date: "`r Sys.Date()`"
package: tidySE
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Overview of the tidySE package}
  %\usepackage[UTF-8]{inputenc}
---

<!-- badges: start -->
[![Lifecycle:maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing) [![R build status](https://github.com/stemangiola/tidySE/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/stemangiola/tidySE/actions)
<!-- badges: end -->

```{r echo=FALSE}
knitr::opts_chunk$set( fig.path = "man/figures/")
```

**Brings SummarizedExperiment to the tidyverse!** 

website: [stemangiola.github.io/tidySE/](https://stemangiola.github.io/tidySE/)

Please have a look also to 

- [tidySCE](https://stemangiola.github.io/tidySCE/) for tidy manipulation of Seurat objects
- [tidyseurat](https://stemangiola.github.io/tidyseurat/) for tidy manipulation of Seurat objects
- [tidybulk](https://stemangiola.github.io/tidybulk/) for tidy high-level data analysis and manipulation 
- [nanny](https://github.com/stemangiola/nanny) for tidy high-level data analysis and manipulation 
- [tidygate](https://github.com/stemangiola/tidygate) for adding custom gate information to your tibble 
- [tidyHeatmap](https://stemangiola.github.io/tidyHeatmap/) for heatmaps produced with tidy principles

<!---

[![Build Status](https://travis-ci.org/stemangiola/tidySE.svg?branch=master)](https://travis-ci.org/stemangiola/tidySE) [![Coverage Status](https://coveralls.io/repos/github/stemangiola/tidySE/badge.svg?branch=master)](https://coveralls.io/github/stemangiola/tidySE?branch=master)

-->

```{r, echo=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


## Functions/utilities available

SummarizedExperiment-compatible Functions | Description
------------ | -------------
`all` | After all `tidySE` is a SummarizedExperiment object, just better

tidyverse Packages | Description
------------ | -------------
`dplyr` | All `dplyr` APIs like for any tibble
`tidyr` | All `tidyr` APIs like for any tibble
`ggplot2` | `ggplot` like for any tibble
`plotly` | `plot_ly` like for any tibble

Utilities | Description
------------ | -------------
`tidy` | Add `tidySE` invisible layer over a SummarizedExperiment object
`as_tibble` | Convert cell-wise information to a `tbl_df`

## Installation

From Github
```{r, eval=FALSE}
devtools::install_github("stemangiola/tidySE")
```

Load libraries used in the examples.

```{r}
library(ggplot2)
library(tidySE)
```


## Create `tidySE`, the best of both worlds!
This is a SummarizedExperiment object but it is evaluated as tibble. So it is fully compatible both with SummarizedExperiment and tidyverse APIs. 

```{r}
pasilla_tidy = tidySE::pasilla %>% tidySE::tidy()
```

**It looks like a tibble**

```{r}
pasilla_tidy
```

**But it is a SummarizedExperiment object after all**

```{r}
pasilla_tidy@assays
```

## Annotation polishing 
We may want to alter the content in a column, for example, shorten single_end to single.

```{r}
pasilla_polished =
  pasilla_tidy %>%
  mutate(type = gsub("_end", "", type)) 

pasilla_polished 
```

## Preliminary plots

```{r}
my_theme <- 	
  list(
    scale_fill_brewer(palette="Set1"),
    scale_color_brewer(palette="Set1") ,
    theme_bw() +
  	theme(
  		panel.border = element_blank(),
  		axis.line = element_line(),
  		panel.grid.major = element_line(size = 0.2),
  		panel.grid.minor = element_line(size = 0.1),
  		text = element_text(size=12),
  		legend.position="bottom",
  		aspect.ratio=1,
  		strip.background = element_blank(),
  		axis.title.x  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10)),
  		axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
  	)
  )
```

We can treat `pasilla_polished` effectively as a normal tibble for plotting. 

Here we plot the distribution of counts per sample.

```{r plot1}
pasilla_polished %>%
	tidySE::ggplot(aes(counts + 1, group=sample, color=`type`)) +
	geom_density() +
	scale_x_log10() +
	my_theme
```
